### 2022-04-11 doc/dd/c3-eval.tex
Ready for review

\begin{tikzpicture}
\definecolor{neutralcolor}{RGB}{0,0,0}
\definecolor{pathcolor}{RGB}{0,0,0}
\definecolor{fillcolor}{RGB}{255,255,255}
\definecolor{background}{RGB}{225,225,225}
\definecolor{highcolor}{RGB}{0,0,0}
\definecolor{lowcolor}{RGB}{0,0,0}
\draw (1.32,6.80) [thin,neutralcolor] -- (0.26,5.74);
\draw (1.32,6.80) [thin,neutralcolor] -- (1.68,5.03);
\draw (1.68,5.03) [thin,neutralcolor] -- (2.74,3.97);
\draw (1.68,5.03) [thin,neutralcolor] -- (1.32,3.27);
\draw (1.32,3.27) [thin,neutralcolor] -- (0.26,2.21);
\draw (1.32,3.27) [thin,neutralcolor] -- (1.68,1.50);
\draw (1.68,1.50) [thin,neutralcolor] -- (2.74,0.44);
\draw (1.68,1.50) [thin,neutralcolor] -- (0.62,0.44);
\draw [thin,fill=fillcolor,draw=fillcolor] (1.68,5.03) circle [radius=0.26];
\node at (1.68,5.03) {{\texttt{*}}};
\draw [thin,fill=fillcolor,draw=fillcolor] (1.68,1.50) circle [radius=0.26];
\node at (1.68,1.50) {{\texttt{*}}};
\draw [thin,fill=fillcolor,draw=fillcolor] (2.74,0.44) circle [radius=0.44];
\node at (2.74,0.44) {$1-w_2$};
\draw [thin,fill=fillcolor,draw=fillcolor] (1.32,6.80) circle [radius=0.26];
\node at (1.32,6.80) {{\texttt{+}}};
\draw [thin,fill=fillcolor,draw=fillcolor] (1.32,3.27) circle [radius=0.26];
\node at (1.32,3.27) {{\texttt{+}}};
\draw [thin,fill=fillcolor,draw=fillcolor] (2.74,3.97) circle [radius=0.44];
\node at (2.74,3.97) {$1-w_1$};
\draw [thin,fill=fillcolor,draw=fillcolor] (0.26,2.21) circle [radius=0.26];
\node at (0.26,2.21) {$w_2$};
\draw [thin,fill=fillcolor,draw=fillcolor] (0.62,0.44) circle [radius=0.26];
\node at (0.62,0.44) {$w_3$};
\draw [thin,fill=fillcolor,draw=fillcolor] (0.26,5.74) circle [radius=0.26];
\node at (0.26,5.74) {$w_1$};
\node [right] at (1.59,6.80) {$s_7$};
\node [left] at (1.41,5.03) {$p_6$};
\node [right] at (1.59,3.27) {$s_5$};
\node [left] at (1.41,1.50) {$p_4$};
\end{tikzpicture}%%