### 2024-01-07 doc/JAIR/lean-section.tex
feat: Lean section

%% Separate file for Lean presentation

\section{A Formally Verified Toolchain}
\label{sect:formally-verified-toolchain}
\label{sect:lean:subtle-condition}

We set out to formally verify the system with two goals in mind:
first, to ensure that the CPOG framework is mathematically sound;
and second, to implement correct-by-construction proof checking
and ring evaluation (the ``Trusted Code'' components of \ref{fig:chain}).
These two goals are achieved with a single proof development
in the \lean{} programming language~\cite{demoura:cade:2021}.
Verification was greatly aided
by the Aesop~\cite{23limperg_aesop_white_box_best_first_proof_search_lean}
automated proof search tactic.
\lean{} is based on a logical foundation
in which expressions have a computational interpretation.
As in other proof assistants such as Isabelle and Coq,
functions defined in the formal system
can be compiled to machine code.
At the same time,
we can state and prove claims about them within the same system,
thereby verifying that our functions compute the intended results.
In this section,
we describe the functionality we implemented,
what we proved about it,
and the assumptions we made.

\paragraph{Data structures and mathematical model.}
When thinking about formal verification,
it is helpful to distinguish between data structures
that play a role in the code being executed,
and \emph{ghost} definitions that serve as a mathematical model,
allowing us to state and prove specifications,
but are erased during compilation and not executed.
We generally store definitions in the two classes
under {\tt Data/} and {\tt Model/},
respectively.

Among the former is our representation of CNF formulas.
Following the DIMACS CNF convention,
a variable is represented as a positive natural number,
a literal is a non-zero integer,
a clause is an array of literals,
and a CNF formula is an array of clauses.
\begin{lstlisting}
def Var := { x : Nat // 0 < x }
def ILit := { i : Int // i ≠ 0 }
abbrev IClause := Array ILit
abbrev ICnf := Array IClause
\end{lstlisting}

A POG is represented as a flat array of elements.
Each element {\tt PogElt} of a POG is either a variable,
a binary disjunction (sum),
or an arbitrary conjunction (product).
\begin{lstlisting}
inductive PogElt where
  | var (x : Var) : PogElt
  | disj (x : Var) (l r : ILit) : PogElt
  | conj (x : Var) (args : Array ILit) : PogElt
\end{lstlisting}
In the first case,
the argument \lstinline{x} is the index of an input variable;
in disjunctions and conjunctions,
it is an extension variable appearing in the CPOG file.
A \lstinline{Pog} is then an array of \lstinline{PogElt}s that is well-founded
in the sense that each element depends only on prior elements in the array.
Note that representing edges as literals
allows us to negate the arguments to \lstinline{disj} and \lstinline{conj}.

On the mathematical side,
our specifications rely on a general theory of propositional logic
mirroring \ref{sect:logical:foundations}.
The type \lstinline{PropForm} describes the syntax of propositional formulas.
It is generic over the type of variables,
so we instantiate it with numeric variables as \lstinline{PropForm Var}.
Assignments of truth values are taken to be total functions \lstinline{PropAssignment Var := Var → Bool}.
Requiring totality is not a limitation:
instead of talking about two equal,
partial assignments to a subset $X' \subseteq X$ of variables,
we can more conveniently talk about two total assignments that agree on $X'$.
We write \lstinline{σ ⊨ φ} when \lstinline{σ : PropAssignment Var} satisfies \lstinline{φ : PropForm Var}.

Functions \lstinline{ILit.toPropForm},
\lstinline{IClause.toPropForm},
\lstinline{ICnf.toPropForm},
and \lstinline{Pog.toPropForm}
relate data structures to the formulas they encode.
For example, given a literal \lstinline{u},
\lstinline{P.toPropForm u} denotes the interpretation
of the node $\nodeu$ corresponding to \lstinline{u} in the POG \lstinline{P}
as a propositional formula $\phi_\nodeu/\neg\phi_\nodeu$
over the input variables.
It is negated if \lstinline{u} has negative polarity.
Lean provides a convenient ``anonymous projection'' notation
that allows writing \lstinline{P.toPropForm u} instead of \lstinline{Pog.toPropForm P u}
when \lstinline{P} has type \lstinline{Pog},
\lstinline{C.toPropForm} instead of \lstinline{IClause.toPropForm C}
when \lstinline{C} has type \lstinline{IClause},
etc.

In order to reason about composite formulas,
we found it easier to work with propositional formulas modulo logical equivalence,
a structure known in logic as the \emph{Lindenbaum--Tarski algebra},
rather than using \lstinline{PropForm} directly.
Its advantage is that equivalent but not syntactically equal formulas
(such as $x \vee \neg x$ and $\top$)
give rise to equal elements in the algebra,
and equality has a privileged position in proof assistants based on type theory:
equals can be substituted for equals in any context.
In this way, forgetting syntactic detail is helpful.
On the other hand,
using the algebra gives rise to some challenges.
The algebra, called {\tt PropFun}, is defined as a quotient,
with Boolean operations and the entailment relation
lifted from the syntax of formulas to the new type.
It is no longer straightforward to say
when an element of the quotient ``depends'' on a variable
since equivalent formulas can refer to different sets of variables.
Instead, we use a semantic notion of dependence
in which an element $\phi$ of the quotient depends on a variable $x$
if and only if there is a truth assignment that satisfies $\phi$,
but falsifies $\phi$ after $x$ is flipped.
\begin{lstlisting}
/-- The *semantic variables* of `φ`
are those it is sensitive to as a Boolean function.
Unlike `vars`, this set is stable under equivalence of formulas. -/
def semVars (φ : PropFun ν) : Set ν :=
  { x | ∃ (τ : PropAssignment ν), τ ⊨ φ ∧ τ.set x (!τ x) ⊭ φ }
\end{lstlisting}

\paragraph{Proof checking.}
The goal of a CPOG proof is to construct a POG
that is equivalent to the input CNF $\inputformula$.
The database of active clauses,
the POG being constructed,
and its root literal,
are stored in a checker state structure {\tt PreState}.
The checker begins by parsing the input formula,
initializing the active clauses to $\theta \leftarrow \inputformula$,
and initializing the POG $P$ to an empty one.
It then processes every step of the CPOG proof,
either modifying its state by adding/deleting clauses in $\theta$
and adding nodes to $P$,
or throwing an exception if a step is incorrect.
Afterwards, it carries out the \textsc{final conditions} check of \ref{subsection:semantics}.

Throughout the process,
we maintain invariants needed to establish the final result.
These ensure that $P$ is partitioned
and that a successful final check entails the logical equivalence
of $\inputformula$ and $\phi_\noder$,
where $\noder$ is the final POG root (\ref{thm:lean:equiv}).
Formally, we define a type {\tt State}
consisting of those {\tt PreState}s
that satisfy all the invariants.
A {\tt State} is a structure
combining {\tt PreState} fields with additional ones
storing computationally irrelevant \emph{ghost state}
that asserts the invariants.
The fields of \lstinline{st : PreState} include
\lstinline{st.inputCnf} for $\inputformula$,
\lstinline{st.clauseDb} for $\theta$,
and \lstinline{st.pog} for $P$.
We write \lstinline{st.pogDefsForm} for the clausal POG definitions formula $\bigwedge_{\nodeu\in P}\theta_u$,
and \lstinline{st.allVars} for all variables (original and extension) added so far.
For any $\nodeu\in P$,
\lstinline{st.pog.toPropForm u} computes $\phi_\nodeu$.

The first invariant states that assignments to input variables
extend uniquely to extension variables defining the POG nodes.
In the formalization, we split this into extension and uniqueness:
\begin{lstlisting}
/-- Any assignment satisfying φ₁ extends to φ₂ while preserving values on X. -/
def extendsOver (X : Set Var) (φ₁ φ₂ : PropForm Var) :=
  ∀ (σ₁ : PropAssignment Var), σ₁ ⊨ φ₁ → ∃ σ₂, σ₁.agreeOn X σ₂ ∧ σ₂ ⊨ φ₂
/-- Assignments satisfying φ are determined on Y by their values on X. -/
def uniqueExt (X Y : Set Var) (φ : PropForm Var) :=
  ∀ (σ₁ σ₂ : PropAssignment Var), σ₁ ⊨ φ → σ₂ ⊨ φ → σ₁.agreeOn X σ₂ →
    σ₁.agreeOn Y σ₂

invariants.extends_pogDefsForm : extendsOver st.inputCnf.vars ⊤ st.pogDefsForm
invariants.uep_pogDefsForm : uniqueExt st.inputCnf.vars st.allVars st.pogDefsForm
\end{lstlisting}
Note that in the definition of \lstinline{uniqueExt}, the arrows associate to the right,
so the definition says that the three assumptions imply the conclusion.
The next invariant guarantees that the set of solutions over the input variables is preserved:
\begin{lstlisting}
def equivalentOver (X : Set Var) (φ₁ φ₂ : PropForm Var) :=
  extendsOver X φ₁ φ₂ ∧ extendsOver X φ₂ φ₁

invariants.equivInput : equivalentOver st.inputCnf.vars st.inputCnf st.clauseDb
\end{lstlisting}
Finally, for every node $\nodeu\in P$ with corresponding literal $u$ we ensure that $\phi_\nodeu$ is partitioned (\ref{def:partitioned-operation-formula}) and relate $\phi_\nodeu$ to its clausal encoding $\theta_u \doteq u \wedge \bigwedge_{\nodev\in P}\theta_v$:
\begin{lstlisting}
def partitioned : PropForm Var → Prop
  | tr | fls | var _ => True
  | neg φ    => φ.partitioned
  | disj φ ψ => φ.partitioned ∧ ψ.partitioned ∧ ∀ τ, ¬(τ ⊨ φ ∧ τ ⊨ ψ)
  | conj φ ψ => φ.partitioned ∧ ψ.partitioned ∧ φ.vars ∩ ψ.vars = ∅

invariants.partitioned : ∀ (u : ILit), (st.pog.toPropForm u).partitioned
invariants.equivalent_lits : ∀ (u : ILit), equivalentOver st.inputCnf.vars
    (u ∧ st.pogDefsForm) (st.pog.toPropForm x)
\end{lstlisting}

The bulk of our work involved showing
that these invariants are indeed maintained by the checker
when going through a valid CPOG proof,
modifying the active clause database and the POG.
Together with additional, technical invariants
about the correctness of cached computations,
they imply the soundness theorem for $P$ with root node $\noder$:

\begin{thm}
\label{thm:lean:equiv}
If the proof checker has assembled POG $P$ with root node $\noder$ starting from input formula $\inputformula$, and \textsc{final conditions} (as stated in Section~\ref{subsection:semantics}) hold of the checker state, then $\inputformula$ is logically equivalent to $\phi_\noder$.
\end{thm}
\begin{proof}
\textsc{Final conditions} imply that the active clausal formula $\theta$ is exactly $\pogformula \doteq \{\{r\}\} \cup \; \bigcup_{\nodeu \in P} \theta_{u}$. The conclusion follows from this and the checker invariants. The full proof is formally verified in Lean.
\end{proof}
After certifying a CPOG proof, the checker can pass its in-memory POG representation to the ring evaluator, along with the partitioning guarantee provided by \texttt{invariants.partitioned}.

\paragraph{Ring evaluation.} We formalized the central quantity (\ref{eqn:rep}) in the ring evaluation problem
(\ref{def:ring_evaluation}) in a commutative ring \lstinline{R} as follows:
\begin{lstlisting}
def weightSum {R : Type} [CommRing R]
    (weight : Var → R) (φ : PropForm Var) (s : Finset Var) : R :=
  ∑ τ in models φ s, ∏ x in s, if τ x then weight x else 1 - weight x
\end{lstlisting}
The rules for efficient ring evaluation of partitioned formulas are expressed as:
\begin{lstlisting}
def ringEval (weight : Var → R) : PropForm Var → R
  | tr       => 1
  | fls      => 0
  | var x    => weight x
  | neg φ    => 1 - ringEval weight φ
  | disj φ ψ => ringEval weight φ + ringEval weight ψ
  | conj φ ψ => ringEval weight φ * ringEval weight ψ
\end{lstlisting}
\ref{prop:ring:eval} is then formalized as follows:
\begin{lstlisting}
theorem ringEval_eq_weightSum (weight : Var → R) {φ : PropForm Var} :
    partitioned φ → ringEval weight φ = weightSum weight φ (vars φ)
\end{lstlisting}
To efficiently compute the ring evaluation of a formula represented by a POG node, we implemented
\lstinline{Pog.ringEval} and then proved that it matches the specification above:
\begin{lstlisting}
theorem ringEval_eq_ringEval (pog : Pog) (weight : Var → R) (x : Var) :
  pog.ringEval weight x = (pog.toPropForm x).ringEval weight
\end{lstlisting}
Applying this to the output of our verified CPOG proof checker, which we know to be partitioned
and equivalent to the input formula $\inputformula$, we obtain a proof that our toolchain computes
the correct ring evaluation of $\inputformula$.

\paragraph{Model counting.} Finally, we established that ring evaluation with the appropriate weights
corresponds to the standard model count. To do so, we defined a function that
carries out an integer calculation of the number of models over a set of variables
of cardinality \lstinline{nVars}:
\begin{lstlisting}
def countModels (nVars : Nat) : PropForm Var → Nat
  | tr       => 2^nVars
  | fls      => 0
  | var _    => 2^(nVars - 1)
  | neg φ    => 2^nVars - countModels nVars φ
  | disj φ ψ => countModels nVars φ + countModels nVars ψ
  | conj φ ψ => countModels nVars φ * countModels nVars ψ / 2^nVars
\end{lstlisting}
We then formally proved that for a partitioned formula whose variables are among a finite set
\lstinline{s}, this computation really does count the number of models over \lstinline{s}:
\begin{lstlisting}
theorem countModels_eq_card_models {φ : PropForm Var} {s : Finset Var} :
  vars φ ⊆ s → partitioned φ → countModels (card s) φ = card (models φ s)
\end{lstlisting}
In particular, taking \lstinline{s} to be exactly the variables appearing in \lstinline{φ},
we have that the number of models of \lstinline{φ} over its variables is
\lstinline{countModels φ (card (vars φ))}.

\subsection{Trust}
In this subsection, we clarify what has been verified and what has to be trusted.
Recall that our first step is to parse CNF and CPOG files in order to read in the initial
formula and the proof. We do not verify this step. Instead, the verified checker exposes
flags \verb+--print-cnf+ and \verb+--print-cpog+ which reprint the consumed formula or proof,
respectively. Comparing this to the actual files using {\tt diff} provides an easy way of ensuring
that what was parsed matches their contents. This involves trusting only the correctness of the
print procedure and {\tt diff}. Similarly, if one wants to establish the correctness of the POG
contained in the CPOG file, one can print out the POG that is constructed by the checker and compare.

Lean's code extraction replaces calculations on natural numbers and integers with
efficient but unverified arbitrary precision versions.
Lean also uses an efficient implementation of arrays; within the
formal system, these are defined in terms of lists, but code extraction replaces them
with dynamic arrays and uses reference counting to allow destructive updates when it is safe
to do so \cite{Ullrich:de:Moura:19}.
Finally, Lean's standard library implements hashmaps in terms of arrays.
Many of the basic properties of hashmaps have been formally verified, but not all.
In particular, we make use of a fold operation whose verification is not yet complete.
Thus our proofs depend on the assumption that the fold operation has the
expected properties.

In summary, in addition to trusting Lean's foundation and kernel checker,
we also have to trust that code extraction respects that foundation,
that the implementation of the fold operation for hashmaps satisfies its description,
and that, after parsing, the computation uses the correct input formula.
All of our specifications have been completely proven and verified relative to these assumptions.

%P: clause db
% We define clause databases in {\tt ClauseDb} as hashmaps
% that store each clause together with a flag
% indicating whether it has been deleted.

%P: files and directories
% Our code is contained in the directory {\tt ProofChecker}.
% The main checker loop is implemented in {\tt Checker/CheckerCore.lean}.
% The ring evaluator and model counter are contained in the
% {\tt Count} directory.